### 2022-04-21 tikz/problem_splitting.tex
almost finished

% Tikz File 'topology_general.tex'
\documentclass{standalone}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,amsmath,amssymb}
\usetikzlibrary{arrows,snakes,backgrounds}
    \begin{document}
    \begin{tikzpicture}[->,shorten >=1pt,auto,semithick]
        \tikzstyle{admmnode}=[
            circle,
            draw=black,
            thick,
            fill=gray,
            text=black
        ]

        \node[admmnode, fill=OliveGreen!40] (A) at (0,1.5){\(1\)};
        \node[admmnode, fill=MidnightBlue!40] (B) at (1.5,0){\(2\)};
        \node[admmnode, fill=BurntOrange!40] (C) at (0,-1.5){\(3\)};
        \node[admmnode, fill=Maroon!40] (D) at (-1.299038105676658,-0.75){\(4\)};
        \node[admmnode, fill=SkyBlue!40] (E) at (-1.299038105676658,0.75){\(5\)};

        \draw[-stealth, line width=1.25] (A) to[bend left=20] node[]{}  (B);
        \draw[-stealth, line width=1.25] (A) to[bend left=0] node[]{} (C);
        \draw[-stealth, line width=1.25] (B) to[bend left=20] node[]{} (C);
        \draw[-stealth, line width=1.25] (C) to[bend left=10] node[]{} (D);
        \draw[-stealth, line width=1.25] (D) to[bend left=10] node[]{} (E);
        \draw[-stealth, line width=1.25] (E) to[bend left=10] node[]{} (A);

    \end{tikzpicture}
\end{document}