### 2022-05-06 tikz/topology_details.tex

% Tikz File 'topology_details.tex'
\documentclass{standalone}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,amsmath,amssymb}
\input{../preamble.tex}
\usetikzlibrary{arrows,snakes,backgrounds}
\begin{document}
    \begin{tikzpicture}[->,shorten >=1pt,auto,semithick, scale=1]
        \tikzstyle{admmnode}=[
            circle,
            draw=black,
            thick,
            text=black,
            minimum size=0.7cm
        ]

        \node[] (C)  at (-2.2,0){\(\cdots\)};

        \node[admmnode, fill=gray!15] (A) at (-1.5,0){\(i\)};
        % \node[draw,text centered] at (-3,-1){\(\mathcal{T}_i = \{\ldots,j,\ldots\}\)};

        \node[admmnode, fill=gray!15] (B) at (1.5,0){\(j\)};
        % \node[draw,text centered] at (3,-1){\(\mathcal{R}_j = \{\ldots,i,\ldots\}\)};

        \node[] (D)  at (2.2,0){\(\cdots\)};

        \draw[-stealth,black, line width=1.2] (A) to[bend left=20] node[above, pos=0.5, sloped, scale=1]{\(\x_i, \hf_i, \{\|\bar{\hf}_1\|^2,\ldots,\|\bar{\hf}_i\|^2\}\)}  (B);

        \draw[stealth-,gray, line width=1.2] (A) to[bend right=20] node[below, pos=0.5, sloped,scale=1]{\((\wf_j)_i,(\uuf_j)_i,\{\|\bar{\hf}_1\|^2,\ldots,\|\bar{\hf}_M\|^2\}\)}  (B);

    \end{tikzpicture}
\end{document}