### 2022-05-06 tikz/topology_inset_simulation.tex

% Tikz File 'topology_general.tex'
\documentclass{standalone}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz,amsmath,amssymb}
\usetikzlibrary{arrows,snakes,backgrounds}
    \begin{document}
    \begin{tikzpicture}[->,shorten >=1pt,auto,semithick]
        \tikzstyle{admmnode}=[
            circle,
            draw=black,
            thick,
            fill=gray,
            text=black
        ]
        \tikzstyle{centernode}=[
            diamond,
            draw=black,
            thick,
            fill=gray!40,
            text=black
        ]

        \node[admmnode, fill=gray!15] (A) at (-1,1){\(1\)};
        \node[admmnode, fill=gray!15] (B) at (1,1){\(2\)};
        \node[admmnode, fill=gray!15] (C) at (1,-1){\(3\)};
        \node[admmnode, fill=gray!15] (D) at (-1,-1){\(4\)};

        \draw[-stealth, line width=1.25] (A) to[bend left=10] node[]{} (B);
        \draw[dashed,-stealth, line width=1.25] (A) to[bend left=10] node[]{} (C);
        \draw[dashed,-stealth, line width=1.25] (A) to[bend left=10] node[]{} (D);
        \draw[-stealth, line width=1.25] (B) to[bend left=10] node[]{} (C);
        \draw[dashed,-stealth, line width=1.25] (B) to[bend left=10] node[]{} (A);
        \draw[dashed,-stealth, line width=1.25] (B) to[bend left=10] node[]{} (D);
        \draw[-stealth, line width=1.25] (C) to[bend left=10] node[]{} (D);
        \draw[dashed,-stealth, line width=1.25] (C) to[bend left=10] node[]{} (A);
        \draw[dashed,-stealth, line width=1.25] (C) to[bend left=10] node[]{} (B);
        \draw[-stealth, line width=1.25] (D) to[bend left=10] node[]{} (A);
        \draw[dashed,-stealth, line width=1.25] (D) to[bend left=10] node[]{} (B);
        \draw[dashed,-stealth, line width=1.25] (D) to[bend left=10] node[]{} (C);

    \end{tikzpicture}
\end{document}